---
title: "Final Paper Work"
output: pdf_document
date: "2022-12-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rtweet)
library(twitteR)
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(tidytext)
library(gridExtra)
library(gtrendsR)
```

### OZ
# extracting the last 1000 tweets from Oz's twitter account
```{r}
oz_tweets <- get_timeline(user = "DrOz", n=1000)

oz_tweets_df <- tbl_df(as.data.frame(oz_tweets))
oz_tweets_df
```

# getting mean of oz tweets for cuttoff point
```{r}
mean(oz_tweets_df$display_text_range)
```
# cleaning oz tweet data 
# separating the time of tweets into year, month, day, hour, minute, second
# splitting number of characters per tweet in half at its mean
```{r}
 oz_tweets_df_timed <- oz_tweets_df %>%
  separate(created_at, c("year", "month", "day"), sep = "-") %>%
  separate(day, c("day", "time"), sep = " ") %>%
  separate(time, c("hour", "minute", "second"), sep = ":") %>%
  mutate(display_text_range_h = display_text_range > 147) %>%
  mutate(display_text_range_l = display_text_range < 147) %>%
  select(favorite_count, retweet_count, year, month, day, hour, display_text_range, display_text_range_h, display_text_range_l) 
```
```{r}
oz_tweets_df_timed
```

# plotting oz tweets over hour of the day by number of favorites, separated by number of characters, cuttoff point = 147 (mean)
```{r}
oz_tweets_df_timed %>%
  select(hour, favorite_count, display_text_range_h) %>%
  ggplot(aes(hour, favorite_count, color = display_text_range_h)) +
  geom_point()+
  labs(x = "Hour of the Day (EST)", y = "Number of Favorites", color = "Above 147 Characters")
```




### OZ END






### FETTERMAN
# extracting the last 1000 fetterman tweets
```{r}
fett_tweets <- get_timeline(user = "JohnFetterman", n=1000)

fett_tweets_df <- tbl_df(as.data.frame(fett_tweets))
fett_tweets_df
```
# getting mean of fetterman tweets for cuttoff point
```{r}
mean(fett_tweets_df$display_text_range)
```
# cleaning fetterman tweet data 
# separating the time of tweets into year, month, day, hour, minute, second
# splitting number of characters per tweet in half at its mean
```{r}
 fett_tweets_df_timed <- fett_tweets_df %>%
  separate(created_at, c("year", "month", "day"), sep = "-") %>%
  separate(day, c("day", "time"), sep = " ") %>%
  separate(time, c("hour", "minute", "second"), sep = ":") %>%
  mutate(display_text_range_h = display_text_range > 137) %>%
  mutate(display_text_range_l = display_text_range < 137) %>%
  select(favorite_count, retweet_count, year, month, day, hour, display_text_range, display_text_range_h, display_text_range_l) 
```
```{r}
fett_tweets_df_timed
```
# plotting fetterman tweets over hour of the day by number of favorites, separated by number of characters, cuttoff point = 137 (mean)
```{r}
fett_tweets_df_timed %>%
  select(hour, favorite_count, display_text_range_h) %>%
  ggplot(aes(hour, favorite_count, color = display_text_range_h)) +
  geom_point()+
  labs(x = "Hour of the Day (EST)", y = "Number of Favorites", color = "Above 137 Characters")
```
# this has a crazy outlier here of more than 600,000 favorites, so I will remove it to make the rest of the data easier to view. I will still consider the data point in my analysis
# removing the outlier - will make a note in analysis
```{r}
fett_tweets_df_timed_outlierfix <- fett_tweets_df_timed %>%
  arrange(desc(favorite_count)) 
  
fett_tweets_df_timed_outlierfix <- fett_tweets_df_timed_outlierfix [-1,] 
fett_tweets_df_timed_outlierfix
  
```
# 
```{r}
fett_tweets_df_timed_outlierfix %>%
  select(hour, favorite_count, display_text_range_h) %>%
  ggplot(aes(hour, favorite_count, color = display_text_range_h)) +
  geom_point()+
  labs(x = "Hour of the Day (EST)", y = "Number of Favorites", color = "Above 147 Characters")
```


### FETTERMAN END

### Pictures/links in tweets
# oz picture/links tweets
```{r}
oz_tweets_picture <- oz_tweets_df %>% 
  filter(!str_detect(text, '^"')) %>%
  count(picture = ifelse(str_detect(text, "t.co"),
                         "Oz Picture/link", "Oz No picture/link"))%>%
  mutate(pictureOz = picture) %>%
  mutate(pictureFett = picture)
oz_tweets_picture
```
# fetterman picture links/tweets
```{r}
fett_tweets_picture <- fett_tweets_df %>% 
  filter(!str_detect(text, '^"')) %>%
  count(picture = ifelse(str_detect(text, "t.co"),
                         "Fetterman Picture/link", "Fetterman No picture/link")) %>%
  mutate(pictureFett = picture) %>%
  mutate(pictureOz = picture)
fett_tweets_picture
```
# joining both to have 1 picture/links tibble
```{r}
oz_fett_tweets_picture <- full_join(oz_tweets_picture, fett_tweets_picture) 
oz_fett_tweets_picture
```
```{r}
ggplot(oz_fett_tweets_picture, aes(picture, n, fill = picture)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "", y = "Number of tweets", fill = "")+
  theme(axis.text.x = element_blank())
```

### Sentiment analysis
# Oz
```{r}
oz_tweet_text <- "([^A-Za-z\\d#@']|'(?![A-Za-z\\d#@]))"
oz_tweet_words <- oz_tweets_df %>%
  filter(!str_detect(text, '^"')) %>%
  mutate(text = str_replace_all(text, "https://t.co/[A-Za-z\\d]+|&amp;", "")) %>%
  unnest_tokens(word, text, token = "regex", pattern = oz_tweet_text) %>%
  filter(!word %in% stop_words$word,
         str_detect(word, "[a-z]"))

oz_tweet_words_r <-oz_tweet_words %>%
  count(word) %>%
  arrange(desc(n))
oz_tweet_words_r 
```
# oz most used words barplot
```{r}
oz_tweet_words_top20 <- oz_tweet_words_r %>%
  top_n(20) %>%
  arrange(desc(n))
oz_tweet_words_top20
ggplot(oz_tweet_words_top15, aes(x=n, y=word))+
  geom_bar(stat = "identity")
```

# Fetterman
```{r}
fett_tweet_text <- "([^A-Za-z\\d#@']|'(?![A-Za-z\\d#@]))"
fett_tweet_words <- fett_tweets_df %>%
  filter(!str_detect(text, '^"')) %>%
  mutate(text = str_replace_all(text, "https://t.co/[A-Za-z\\d]+|&amp;", "")) %>%
  unnest_tokens(word, text, token = "regex", pattern = fett_tweet_text) %>%
  filter(!word %in% stop_words$word,
         str_detect(word, "[a-z]"))

fett_tweet_words_r <-fett_tweet_words %>%
  count(word) %>%
  arrange(desc(n))
fett_tweet_words_r 
```
#fetterman most used words barplot
```{r}
fett_tweet_words_top15 <- fett_tweet_words_r %>%
  top_n(20) %>%
  arrange(desc(n))

ggplot(fett_tweet_words_top15, aes(x=n, y=word))+
  geom_bar(stat = "identity")
```

# both
# oz sentiment of tweets
```{r}
sentiments

oz_tweet_words_sentiment <- oz_tweet_words_r %>%
  inner_join(sentiments, by = "word") %>%
  select(sentiment, word) %>%
  count(sentiment) 
oz_tweet_words_sentiment
```
# fetterman sentiment of tweets
```{r}
fett_tweet_words_sentiment <- fett_tweet_words_r %>%
  inner_join(sentiments, by = "word") %>%
  select(sentiment, word) %>%
  count(sentiment) 
fett_tweet_words_sentiment
```
# sentiment plots for both twitter accounts - oz on left, fetterman on right
```{r}
oz_sentiment_plot <- ggplot(oz_tweet_words_sentiment, aes(sentiment, n))+
  geom_bar(stat = "identity", position = "dodge")

fett_sentiment_plot <- ggplot(fett_tweet_words_sentiment, aes(sentiment, n, ylim=c(0,250)))+
  geom_bar(stat = "identity", position = "dodge")

grid.arrange(oz_sentiment_plot, fett_sentiment_plot, ncol=2)
```
####


# constituency analysis
```{r}
names <- gtrends(c("dr", "oz", "john", "fetterman"), 
               geo = "US-PA", time= "2022-06-20 2022-11-9", onlyInterest = TRUE)
plot(names)
```

# oz twitter account's most used words
# 'Fetterman' (1), 'John' (2), 'Pennsylvania' (3), Rt(5), Pa(6), 'Pennsylvanians' (9), 'Oz' (17) - *'Word' (rank of most used on Oz's account)* - I have removed these words from my gtrends search as I have deemed them not useful to my analysis. 'Policies' (8) has been changed to 'policy' for my search. 
```{r}
google_oz_1t5 <- gtrends(c("crime", "biden", "policy", 'radical', 'washington'), 
               geo = "US-PA", time= "2022-06-20 2022-11-9", onlyInterest = TRUE)

google_oz_6t10 <- gtrends(c("debate", 'senate', 'change', 'inflation', 'energy'), 
               geo = "US-PA", time= "2022-06-20 2022-11-9", onlyInterest = TRUE)
google_oz_1t5_plot <- plot(google_oz_1t5)
google_oz_6t10_plot <- plot(google_oz_6t10)
```


#GTRENDS FETTERMAN
# gtrends search by fetterman's twitter account's most used words 
```{r}
google_fett_1t5 <- gtrends(c("vote", "people", "senate", 'campaign', 'abortion'), 
               geo = "US-PA", time= "2022-06-20 2022-11-9", onlyInterest = TRUE)

google_fett_6t10 <- gtrends(c("fight", 'election', 'union', 'life', 'philly'), 
               geo = "US-PA", time= "2022-06-20 2022-11-9", onlyInterest = TRUE)
google_fett_1t5_plot <- plot(google_fett_1t5)
google_fett_6t10_plot <- plot(google_fett_6t10)
```

# PLOT GTRENDS BOTH
# plots of google searches over time for both's twitter's favorite words
```{r}
grid.arrange(google_oz_1t5_plot, google_oz_6t10_plot, ncol=1) 
grid.arrange(google_fett_1t5_plot, google_fett_6t10_plot, ncol=1)
```
# OZ GTRENDS facets
# plots of google searches for oz's twitter's favorite words 1-5
```{r}
google_oz_1t5_df <- as.data.frame(google_oz_1t5)
google_oz_1t5_df

google_oz_1t5_df$interest_over_time.hits <- as.numeric(google_oz_1t5_df$interest_over_time.hits)

google_oz_1t5_df %>%
  qplot(x = interest_over_time.date, y = interest_over_time.hits, data = ., 
        geom = "point", facets = . ~ interest_over_time.keyword)
```
# plots of google searches for oz's twitter's favorite words 6-10
```{r}
google_oz_6t10_df <- as.data.frame(google_oz_6t10)
google_oz_6t10_df

google_oz_6t10_df$interest_over_time.hits <- as.numeric(google_oz_6t10_df$interest_over_time.hits)

google_oz_6t10_df %>%
  qplot(x = interest_over_time.date, y = interest_over_time.hits, data = ., 
        geom = "point", facets = . ~ interest_over_time.keyword)
```

# FETTERMAN GTRENDS facets
# plots of google searches for fetterman's twitter's favorite words 1-5
```{r}
google_fett_1t5_df <- as.data.frame(google_fett_1t5)
google_fett_1t5_df

google_fett_1t5_df$interest_over_time.hits <- as.numeric(google_fett_1t5_df$interest_over_time.hits)

google_fett_1t5_df %>%
  qplot(x = interest_over_time.date, y = interest_over_time.hits, data = ., 
        geom = "point", facets = . ~ interest_over_time.keyword)
```
# plots of google searches for fetterman's twitter's favorite words 6-10
```{r}
google_fett_6t10_df <- as.data.frame(google_fett_6t10)
google_fett_6t10_df

google_fett_6t10_df$interest_over_time.hits <- as.numeric(google_fett_6t10_df$interest_over_time.hits)

google_fett_6t10_df %>%
  qplot(x = interest_over_time.date, y = interest_over_time.hits, data = ., 
        geom = "point", facets = . ~ interest_over_time.keyword)
```







